---
title: "Hospital Length of Stays"
author: "Keith Douglas"
date: "2024-10-15"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

```{r}
library(tidyverse)
library(NHSRdatasets)
library(knitr)
library(kableExtra)
```

# Load the data from the package
```{r}
data("LOS_model")
```

# Inspect
```{r}
str(LOS_model)
head(LOS_model)
```


# Make Death a factor
```{r}
#0 is survived, 1 is died
hospital_data <- LOS_model %>%
  mutate(Death = factor(Death))
hospital_data
#When you print, it says <fct> under Death instead of <int>
```


# Recode Death levels
```{r}
hospital_data <- hospital_data %>%
  mutate(Death = Death %>%
           fct_recode("Survived" = "0",
                      "Died" = "1"))
#The new name must be on the left and old must be on the right
```


# Create a summary table where each combination of Organisation and Death gets a count (n).
```{r}
hospital_data_summary <- hospital_data %>%
  group_by(Organisation, Death) %>%
  tally()
hospital_data_summary
```

# Make a wide table with Dead and Survived as rows with a column for each Trust
```{r}
hospital_data_wide <- hospital_data_summary %>%
  pivot_wider(
    names_from = Organisation,
    values_from = n
  )
hospital_data_wide

#Wider is from
#Longer is to
#Names from = column names
#Organisations are an ordered factor so that's why the pivot_wider function made them in order from Trust 1 in increasing order
```

# Another pivot with Survived and Died as columns, Trusts as rows.
# Also calculate the % survived for each Trust
```{r}
hospital_data_wide_pretty <- hospital_data_summary %>%
  pivot_wider(
    names_from = Death,
    values_from = n
  ) %>%
  mutate(Total = Survived + Died,
         Percent_Survived = (Survived/Total) * 100)
hospital_data_wide_pretty
```


# Make the wide table pretty with kable()
```{r}
hospital_data_wide_pretty %>%
  kable(
    col.names = c("Trust", "Survived", "Died", "Total", "Percent Survived"),
    digits = 0,
    caption = "Hospital Length of Stages data: Percent Survived by Trust",
    align = "lcccc"
  ) %>%
  kable_styling("striped", full_width = FALSE) %>%
  footnote("Data from LOS_model")


hospital_data_wide_pretty
#Digits function rounds the values
#Col.names renames the columns
```


# Let's knit to PDF


